Arrows for Parallel Computations

### 2017-12-19 paper/content/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}
Arrows are a generic concept that allows for powerful composition
combinators. To our knowledge we are first to represent
\emph{parallel} computation with Arrows.
% \olcomment{that strange arrows-based robot interaction paper from 1993 or so! clearly discuss in related work} \done
%
Arrows turn out to be a useful tool for composing parallel
programs. We do not have to introduce new monadic types that wrap the
computation. Instead, we use Arrows in the same manner one uses sequential pure functions. 
%
This work features multiple parallel backends: the already available parallel Haskell flavours. Parallel Arrows, as presented here, feature an implementation of the |ArrowParallel| type class for GpH Haskell, |Par| Monad, and Eden. With our approach parallel programs can be ported across these flavours with little to no effort. It is quite straightforward to add further backends. 
%
%
Performance-wise, Parallel Arrows are on par with existing parallel Haskells, as they do not introduce any notable overhead in our benchmarks.
%
The benefit is, however, the greatly increased portability of parallel programs.

@@

Our PArrows DSL can be expanded to further parallel Haskells. More specifically we target HdpH \cite{Maier:2014:HDS:2775050.2633363} for this future extension. HdpH is a modern distributed Haskell that would benefit from our Arrows notation. Further Future-aware versions of Arrow combinators can be defined. Existing combinators could also be improved.
Arrow-based notation might enable further compiler optimizations.

More experiences with seamless porting of parallel PArrows-based programs across the backends are welcome.
Of course, we are working ourselves on expanding both our skeleton library and the number of skeleton-based parallel programs that use our DSL to be portable across flavours of parallel Haskells.
It would also be interesting to see a hybrid of PArrows and Accelerate \cite{McDonell:2015:TRC:2887747.2804313}.
Ports of our approach to other languages like Frege or Java directly are in an early development stage.

>>>>

\section{Conclusion}
\label{sec:conclusion}
Arrows are a generic concept that allows for powerful composition
combinators. To our knowledge we are first to represent
\emph{parallel} computation with Arrows.
% \olcomment{that strange arrows-based robot interaction paper from 1993 or so! clearly discuss in related work} \done
%
Arrows turn out to be a useful tool for composing parallel
programs. We have shown, that, in order to have a generic and extensible parallel Haskell, we do not have to restrict ourselves to a monadic interface. Instead, we use Arrows which are a more general concept than Monads, but still allow for powerful composition. Furthermore, we think Arrows are a better fit to parallelize pure code than a monadic solution as regular functions are already Arrows and can be used with our DSL in a more natural way. We manage to retain this nice property of parallel Haskells like Eden or GpH (which use pure constructs) while still having a generic and composable interface.
%
The DSL natively allows for parallelization of monadic code via the Kleisli type and additionally allows to parallelize any Arrow type that has an instance for |ArrowChoice| (note that some skeletons require an additional |ArrowLoop| instance).
%
Parallel Arrows, as presented here, feature an implementation of the |ArrowParallel| type class for GpH Haskell, |Par| Monad, and Eden. With our approach parallel programs can be ported across these flavours with little to no effort. It is quite straightforward to add further backends.
%
%
Performance-wise, Parallel Arrows are on par with existing parallel Haskells, as they only introduce minor overhead in some of our benchmarks.
%
The benefit is, however, the greatly increased portability of parallel programs.

@@

Our PArrows DSL can be expanded to further parallel Haskells. More specifically we target HdpH \cite{Maier:2014:HDS:2775050.2633363} for this future extension. HdpH is a modern distributed Haskell that would benefit from our Arrow notation. Further Future-aware versions of Arrow combinators can be defined. Existing combinators could also be improved.
Arrow-based notation might enable further compiler optimizations.

More experiences with seamless porting of parallel PArrows-based programs across the backends are welcome.
Of course, we are ourselves working on expanding both our skeleton library and the number of skeleton-based parallel programs that use our DSL to be portable across flavours of parallel Haskells.
It would also be interesting to see a hybrid of PArrows and Accelerate \cite{McDonell:2015:TRC:2887747.2804313}.
Ports of our approach to other languages like Frege, Eta or Java directly are in an early development stage.
